---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# novahelpers

<!-- badges: start -->
<!-- badges: end -->

Dallas's miscellanious helper functions, namely for summarizing and reporting data.

## Installation

You can install the development version of novahelpers from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("DallasNovakowski/novahelpers")
```

## Usage


```{r example, warning=F, message = F}
library(tidyverse)
library(emmeans)
library(palmerpenguins) # dataset

library(novahelpers)
```

Load data

```{r}
df <- palmerpenguins::penguins

# drop rows with missing values
df <- df[complete.cases(df)==TRUE, ]
```


## make_summary()

Summarizes our relevant descriptive stats

```{r cars}
flipper_summary <- make_summary(data = df, dv = flipper_length_mm, grouping1 = species, grouping2 = sex)

knitr::kable(flipper_summary)
```

Running ANOVA, nothing i take credit for.

```{r}
# basic fit
# Setting contrasts
contrasts(df$species) <- contr.sum

contrasts(df$sex) <- contr.sum

# Type 3 anova with orthogonal contrasts
flipper_fit <- lm(flipper_length_mm ~ species*sex, data = df)

flipper_anova <- car::Anova(flipper_fit, type = 3)

flipper_emmeans <- emmeans::emmeans(flipper_fit, specs = pairwise ~ species*sex)
```


## merge_emmeans_summary()

Combines our `make_summary()` output with `emmeans` object

```{r}
flipper_summary <- merge_emmeans_summary(flipper_summary, flipper_emmeans)
knitr::kable(flipper_summary)
```

## calculate_and_merge_effect_sizes()

Merges raw comparisons and statistical tests with cohen's d effect sizes. Currently only does all tukey-adjusted comparisons

```{r}
flipper_contrasts  <- calculate_and_merge_effect_sizes(flipper_emmeans, flipper_fit)

knitr::kable(flipper_contrasts)
```


```{r}
# Extract effect size (partial eta squared) from anova
flipper_anova_pes <- effectsize::eta_squared(flipper_anova,
                                             alternative="two.sided",
                                             verbose = FALSE)
# Convert anova table into dataframe
flipper_anova <- data.frame(flipper_anova)


flipper_anova <- flipper_anova[!(rownames(flipper_anova) == "(Intercept)"), ]

# import effect size estimates and confidence intervals to anova dataframe
flipper_anova[1:3,"pes_ci95_lo"] <- flipper_anova_pes$CI_low
flipper_anova[1:3,"pes_ci95_hi"] <- flipper_anova_pes$CI_high
flipper_anova[1:3,"pes"] <- flipper_anova_pes$Eta2

# round all numeric columns to 2 decimal places
flipper_anova <- flipper_anova %>%
  dplyr::mutate_if(is.numeric, function(x) round(x, 2))
```


## report_tidy_anova_etaci()

Used for extracting results from anova table into reportable text

```{r}
report_tidy_anova_etaci(flipper_anova, "sex")
```

You can use this inline function inline as well, `r report_tidy_anova_etaci(flipper_anova, "sex")`


## report_tidy_t()


```{r}
flipper_sex_ttest <- t.test(flipper_length_mm ~ sex, data=df) %>%
  report::report() %>%
  data.frame() %>%
  janitor::clean_names()

knitr::kable(flipper_sex_ttest)

report_tidy_t(flipper_sex_ttest, teststat = T)
```

Works on t-tests, `r report_tidy_t(flipper_sex_ttest, teststat = T)`






... and it works on emmmean contrasts from ANOVA, `r report_tidy_t(
                           flipper_contrasts[flipper_contrasts$contrast =="Adelie female - Adelie male",])`

```{r}
report_tidy_t(
                           flipper_contrasts[flipper_contrasts$contrast =="Adelie female - Adelie male",], 
                           italicize = FALSE, 
                           ci = FALSE)
```


# Putting it Together With a Plot, with theme_basic()

```{r, warning=F, message = F}
#Just for this plotting workflow
library(ggpp) # for position_dodge2nudge
library(cowplot)
library(ggdist)
library(ggpubr) # significance brackets
library(ggtext)

# Define color palette
nova_palette <- c("#78AAA9", "#FFDB6E")

ggplot(data = df,
       aes(y = flipper_length_mm, # our dependent/response/outcome variable 
           x = species,  # our grouping/independent/predictor variable
           fill = sex)) +  # our third grouping/independent/interaction variable
    ggdist::stat_slab( alpha = .3,
                     adjust = 2,
                     side = "left", 
                     scale = 0.4, 
                     show.legend = F, 
                     position = position_dodge(width = .6), 
                     .width = c(.50, 1),
                     aes(fill_ramp = after_stat(level))) +
  ## Add stacked dots
  ggdist::stat_dots(alpha = 0.7,
                    side = "left", 
                    scale = 0.4, 
                    aes(color = sex,
                    fill = sex),

                    position = position_dodge(width = .6)) + 
  geom_text(data = flipper_summary, 
            aes(x = species, 
                y = mean, 
                label = round(mean,1)),
            color="black", 
            size = 2, 
            position = position_dodge2nudge(x = .08, width = .6)) +
    geom_pointrange(data = flipper_summary, # our externally-defined summary dataframe
                  aes(x = species,  # our independent variable
                      y = mean, # our outcome/dependent variable
                      ymin = loci,  # lower-bound confidence interval
                      ymax = upci # upper-bound confidence interval
                  ), 
                  show.legend = F,
                  position = position_dodge2nudge(x= -.03, width = .6)) +
  
  
    scale_colour_manual(values = nova_palette, 
                              aesthetics = c("fill", "color")) +   
  ## define amount of fading
ggdist::scale_fill_ramp_discrete(range = c(0.0, 1),
                                   aesthetics = c("fill_ramp")) + 
       labs(subtitle = paste0("**species**: ", report_tidy_anova_etaci(flipper_anova,"species"), "<br>",
                           "**sex**: ", report_tidy_anova_etaci(flipper_anova,"sex"), "<br>",
                           "**species*sex**: ",report_tidy_anova_etaci(flipper_anova,"species:sex"))
         ) + # add a density slab

  
  ggpubr::geom_bracket(inherit.aes = FALSE, # necessary for factorial design
                       tip.length = 0.02, 
                       vjust = -0.6,
                       xmin = 1.12, # You need to play with these by hand
                       xmax = 1.82, 
                       y.position = 210 ,
                       label.size = 2.1,
                       label = paste0(
                         report_tidy_t(
                           flipper_contrasts[flipper_contrasts$contrast =="Chinstrap female - Adelie male",], 
                           italicize = FALSE, 
                           ci = FALSE)) # content of your bracket text
                       
  ) +
  ggpubr::geom_bracket(inherit.aes = FALSE, 
                       tip.length = -0.02, 
                       vjust = 0.3,
                       xmin = 2.12, 
                       xmax = 3.12, 
                       y.position = 192 ,
                       label.size = 2.1,
                       label = paste0(
                         report_tidy_t(
                           flipper_contrasts[flipper_contrasts$contrast =="Chinstrap male - Gentoo male",], 
                           italicize = FALSE, 
                           ci = FALSE)) # content of your bracket text
  ) +

  theme_basic() +
  guides(fill_ramp = "none")
```


