---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# novahelpers

<!-- badges: start -->

<!-- badges: end -->

Dallas's miscellanious helper functions, namely for summarizing and reporting data.

## Installation

You can install the development version of novahelpers from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("DallasNovakowski/novahelpers")
```

## Usage

```{r example, warning=F, message = F}
library(tidyverse)
library(emmeans)
library(palmerpenguins) # dataset

library(novahelpers)
```

Load data

```{r}
df <- palmerpenguins::penguins

# drop rows with missing values
df <- df[complete.cases(df)==TRUE, ]
```

## run_summary()

Summarizes our relevant descriptive stats

```{r cars}
flipper_summary <- run_summary(df, c("sex", "species"), "flipper_length_mm")

knitr::kable(flipper_summary)
```



Running ANOVA, nothing i take credit for.

```{r}
# basic fit
# Setting contrasts
contrasts(df$species) <- contr.sum

contrasts(df$sex) <- contr.sum

# consider just using afex - https://www.rdocumentation.org/packages/afex/versions/1.3-0
#a <- aov_ez("id", "log_rt", fhch, between = "task", within = c("stimulus", "length"))

# Type 3 anova with orthogonal contrasts
flipper_fit <- lm(flipper_length_mm ~ species*sex, data = df)

flipper_anova <- car::Anova(flipper_fit, type = 3)

flipper_emmeans <- emmeans::emmeans(flipper_fit, specs = pairwise ~ species*sex)
```

## merge_emmeans_summary()

Combines our `run_summary()` output with `emmeans` object

```{r}
flipper_summary <- merge_emmeans_summary(flipper_summary, flipper_emmeans)
knitr::kable(flipper_summary)
```

## Planned Comparisons

See <https://aosmith.rbind.io/2019/04/15/custom-contrasts-emmeans/> and <https://cran.r-project.org/web/packages/emmeans/vignettes/comparisons.html#linfcns> for custom contrasts

```{r}
flipper_just_emmeans <- emmeans::emmeans(flipper_fit, specs = ~ species*sex)

# correponds to row numbers in  flipper_just_emmeans
chinstrap_female = c(0, 1, 0, 0, 0, 0)
adelie_female = c(1, 0, 0, 0, 0, 0)

contrast(flipper_just_emmeans, method = list("Adelie female - Chinstrap female" =  adelie_female-chinstrap_female) )
```


## calculate_and_merge_effect_sizes()

Merges raw comparisons and statistical tests with cohen's d effect sizes. Currently only does all tukey-adjusted comparisons

```{r}
flipper_contrasts  <- calculate_and_merge_effect_sizes(flipper_emmeans, flipper_fit)

knitr::kable(flipper_contrasts)
```

```{r}
# Extract effect size (partial eta squared) from anova
flipper_anova_pes <- effectsize::eta_squared(flipper_anova,
                                             alternative="two.sided",
                                             verbose = FALSE)
# Convert anova table into dataframe
flipper_anova <- data.frame(flipper_anova)


flipper_anova <- flipper_anova[!(rownames(flipper_anova) == "(Intercept)"), ]

# import effect size estimates and confidence intervals to anova dataframe
flipper_anova[1:3,"pes_ci95_lo"] <- flipper_anova_pes$CI_low
flipper_anova[1:3,"pes_ci95_hi"] <- flipper_anova_pes$CI_high
flipper_anova[1:3,"pes"] <- flipper_anova_pes$Eta2

# round all numeric columns to 2 decimal places
flipper_anova <- flipper_anova %>%
  dplyr::mutate_if(is.numeric, function(x) round(x, 2))
```

## report_tidy_anova_etaci()

Used for extracting results from anova table into reportable text

```{r}
report_tidy_anova_etaci(flipper_anova, "sex")
```

You can use this inline function inline as well, `r report_tidy_anova_etaci(flipper_anova, "sex")`

## report_tidy_t()

```{r}
flipper_sex_ttest <- t.test(flipper_length_mm ~ sex, data=df) %>%
  report::report() %>%
  data.frame() %>%
  janitor::clean_names()

knitr::kable(flipper_sex_ttest)

report_tidy_t(flipper_sex_ttest, teststat = T)
```

Works on t-tests, `r report_tidy_t(flipper_sex_ttest, teststat = T)`

... and it works on emmmean contrasts from ANOVA, `r report_tidy_t(                            flipper_contrasts[flipper_contrasts$contrast =="Adelie female - Adelie male",])`

```{r}
report_tidy_t(
                           flipper_contrasts[flipper_contrasts$contrast =="Adelie female - Adelie male",], 
                           italicize = FALSE, 
                           ci = FALSE)
```

# Putting it Together With a Plot, with theme_basic()

```{r, warning=F, message = F}
#Just for this plotting workflow
library(ggpp) # for position_dodge2nudge
library(cowplot)
library(ggdist)
library(ggpubr) # significance brackets
library(ggtext)

# Define color palette
nova_palette <- c("#78AAA9", "#FFDB6E")

ggplot(data = df,
       aes(y = flipper_length_mm, # our dependent/response/outcome variable 
           x = species,  # our grouping/independent/predictor variable
           fill = sex)) +  # our third grouping/independent/interaction variable
    ggdist::stat_slab( alpha = .3,
                     adjust = 2,
                     side = "left", 
                     scale = 0.4, 
                     show.legend = F, 
                     position = position_dodge(width = .6), 
                     .width = c(.50, 1),
                     aes(fill_ramp = after_stat(level))) +
  ## Add stacked dots
  ggdist::stat_dots(alpha = 0.7,
                    side = "left", 
                    scale = 0.4, 
                    aes(color = sex,
                    fill = sex),

                    position = position_dodge(width = .6)) + 
  geom_text(data = flipper_summary, 
            aes(x = species, 
                y = mean, 
                label = round(mean,1)),
            color="black", 
            size = 2, 
            position = position_dodge2nudge(x = .08, width = .6)) +
    geom_pointrange(data = flipper_summary, # our externally-defined summary dataframe
                  aes(x = species,  # our independent variable
                      y = mean, # our outcome/dependent variable
                      ymin = loci,  # lower-bound confidence interval
                      ymax = upci # upper-bound confidence interval
                  ), 
                  show.legend = F,
                  position = position_dodge2nudge(x= -.03, width = .6)) +
  
  
    scale_colour_manual(values = nova_palette, 
                              aesthetics = c("fill", "color")) +   
  ## define amount of fading
ggdist::scale_fill_ramp_discrete(range = c(0.0, 1),
                                   aesthetics = c("fill_ramp")) + 
       labs(subtitle = paste0("**species**: ", report_tidy_anova_etaci(flipper_anova,"species"), "<br>",
                           "**sex**: ", report_tidy_anova_etaci(flipper_anova,"sex"), "<br>",
                           "**species*sex**: ",report_tidy_anova_etaci(flipper_anova,"species:sex"))
         ) + # add a density slab

  
  ggpubr::geom_bracket(inherit.aes = FALSE, # necessary for factorial design
                       tip.length = 0.02, 
                       vjust = -0.6,
                       xmin = 1.12, # You need to play with these by hand
                       xmax = 1.82, 
                       y.position = 210 ,
                       label.size = 2.1,
                       label = paste0(
                         report_tidy_t(
                           flipper_contrasts[flipper_contrasts$contrast =="Chinstrap female - Adelie male",], 
                           italicize = FALSE, 
                           ci = FALSE)) # content of your bracket text
                       
  ) +
  ggpubr::geom_bracket(inherit.aes = FALSE, 
                       tip.length = -0.02, 
                       vjust = 0.3,
                       xmin = 2.12, 
                       xmax = 3.12, 
                       y.position = 192 ,
                       label.size = 2.1,
                       label = paste0(
                         report_tidy_t(
                           flipper_contrasts[flipper_contrasts$contrast =="Chinstrap male - Gentoo male",], 
                           italicize = FALSE, 
                           ci = FALSE)) # content of your bracket text
  ) +

  theme_basic() +
  guides(fill_ramp = "none")
```

For posterity, here is a plot with just faded density slabs

```{r}
ggplot(data = df,
       aes(y = flipper_length_mm, # our dependent/response/outcome variable 
           x = species,  # our grouping/independent/predictor variable
           fill = sex)) +  # our third grouping/independent/interaction variable
    ggdist::stat_slab(
                     side = "left", 
                     scale = 0.5, 
                     position = position_dodge(width = .6), 
                     .width = c(.50, 1),
                     aes(fill_ramp = after_stat(level))) +
  ggdist::scale_fill_ramp_discrete(range = c(0.2, 1),
                                   aesthetics = c("fill_ramp")) + 
      scale_colour_manual(values = nova_palette, 
                              aesthetics = c("fill")) +   
  guides(fill_ramp = "none") +

  geom_text(data = flipper_summary, 
            aes(x = species, 
                y = y25 - .1*sd(df$flipper_length_mm), 
                label = round(mean,1)),
            color="black", 
            size = 2.5, 

            position = position_dodge2nudge(x= -.06, width = .6)) +
    geom_pointrange(data = flipper_summary, # our externally-defined summary dataframe
                  aes(x = species,  # our independent variable
                      y = mean, # our outcome/dependent variable
                      ymin = loci,  # lower-bound confidence interval
                      ymax = upci # upper-bound confidence interval
                  ), 
                  show.legend = F,
                  position = position_dodge2nudge(x= -.05, width = .6)) + 
  theme_basic()
```


# many

Often, you may need to run your analysis on multiple variables (e.g., mediators). The `many` family of functions lets you designate multiple dependent variables to run summaries, lm, emmeans, anovas, and merging emmeans/summaries. These functions each return a list corresponding to your dependent variable name.

## run_many_summaries()

This is a useful function for running run_summary() over many variables, and storing them all in one place

```{r}

dv_vars <- c("bill_length_mm", "flipper_length_mm", "body_mass_g")
grouping_vars <- c("species", "sex")

summary_list <- run_many_summaries(data = df,
                  summarization_vars = dv_vars,
                  group_vars = grouping_vars)
summary_list
```


## run_many_lm()

```{r}
lm_results_list <- run_many_lm(data = df, 
                                independent_variable = grouping_vars, 
                                dependent_variables = dv_vars)


# or just run

# results <- list()
# for (var in c("bill_length_mm", "flipper_length_mm", "body_mass_g")) {
#   formula <- as.formula(paste(var, "~", paste(c("species", "sex"), 
#                                               collapse = "*")))
#   model <- lm(formula, data = df)
#   results[[var]] <- model
# }

```

## run_many_emmeans()

```{r}
emmeans_specs <- ~ species * sex


emmeans_results_list <- run_many_emmeans(lm_results_list, emmeans_specs)

emmeans_results_list
```


## run_many_anovas()

```{r}
anova_results_list <- run_many_anovas(lm_results_list, type = "III")

anova_results_list
```



## merge_many_emmeans_summary()

```{r}
summary_list <- merge_many_emmeans_summary(summary_list, 
                           emmeans_results_list, 
                           grouping_vars = c("species", "sex"))

summary_list
```

